---
title: "DA5030.Loan"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Q.1. Build an R Notebook of the bank loan decision tree example in the textbook on pages 136 to 149; the CSV file is available for download below. Show each step and add appropriate documentation. Note that the provided dataset uses values 1 and 2 in default column whereas the book has no and yes in the default column. To fix any problems replace "no" with "1" and "yes" with "2" in the code that for matrix_dimensions. Alternatively, change the line
error_cost <- matrix(c(0, 1, 4, 0), nrow = 2, dimnames = matrix_dimensions) to error_cost <- matrix(c(0, 1, 4, 0), nrow = 2). 

```{r}
install.packages("C50", repos = "http://cran.us.r-project.org")
install.packages("gmodels", repos = "http://cran.us.r-project.org")
```

```{r}
#collecting data
#setwd("C:/Users/omkar/OneDrive/Desktop/R/DA5030.A5")
credit <- read.csv("credit.csv")

#exploring data
str(credit)
table(credit$checking_balance)
table(credit$savings_balance)
summary(credit$months_loan_duration)
summary(credit$amount)
table(credit$default)

set.seed(123)
train_sample <- sample(1000, 900)
str(train_sample)

#splitting data into training and testing dataset
credit_train <- credit[train_sample, ]
credit_test  <- credit[-train_sample, ]

prop.table(table(credit_train$default))
prop.table(table(credit_test$default))


library(C50)
#converting 'default' data from numeric to factor
credit_train$default <- as.factor(credit_train$default)

#training model on data
credit_model <- C5.0(credit_train[-17], credit_train$default)
credit_model
summary(credit_model)

library(gmodels)

#evaluating model
credit_pred <- predict(credit_model, credit_test)
CrossTable(credit_test$default, credit_pred, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE,dnn = c('actual default', 'predicted default')) 

#improving model performance by adaptive boosting
credit_boost10 <- C5.0(credit_train[-17], credit_train$default, trials = 10)
credit_boost10
summary(credit_boost10)
credit_boost_pred10 <- predict(credit_boost10, credit_test) 
CrossTable(credit_test$default, credit_boost_pred10, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, dnn = c('actual default', 'predicted default')) 

```

```{r}

#assigning penalty to costlier errors
matrix_dimensions <- list(c("1", "2"), c("1", "2"))
names(matrix_dimensions) <- c("predicted", "actual")
matrix_dimensions

error_cost <- matrix(c(0, 1, 4, 0), nrow = 2, dimnames = matrix_dimensions)
error_cost

credit_cost <- C5.0(credit_train[-17], credit_train$default, costs = error_cost)
credit_cost_pred <- predict(credit_cost, credit_test)
CrossTable(credit_test$default, credit_cost_pred, prop.chisq = FALSE, prop.c = FALSE, prop.r = FALSE, 
           dnn = c('actual default', 'predicted default'))
```